---
title: 'Lecture 07: Performance'
output: html_document
---


```{r setup, echo = FALSE}
## setup
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", out.width = "50%", echo = TRUE)
invisible(suppressPackageStartupMessages(testthat::expect_equal))
```

In this exercise you should write faster versions of the functions that are given.

# Exercise 0: Test things

Define a variable `zero` with value `0`.
This is to try out the setup.
```{r}
zero <- 0  # change this?
```

# Exercise 1: Sum

Write a function that sums the numbers from 1 to `n`, where `n` is a positive integer.

```{r}
nsum <- function(n) {
   result <- numeric(0)
  for (i in seq_len(n)) result[[length(result) + 1]] <-  i
  sum(result)
}
```

#### Tests

```{r, error = TRUE, echo = FALSE}

testthat::expect_equal(nsum(1), 1)

testthat::expect_equal(nsum(10), 55)

testthat::expect_equal(nsum(100), 5050)

testthat::expect_equal(nsum(73), 2701)

testthat::expect_equal(nsum(69), 2415)

```

## Exercise 2: Number of Missings

Write a function that counts the number of missing values in each row of a `data.frame` `df`.
The return value should be a integer-valued `numeric` vector with length equal to the number of rows of `df`.
The `i`th entry of the return value should indicate the number of missing values in the `i`th row of `df`.
You should also consider the case of having zero-row or zero-column `data.frame`s.

```{r}
countMissings <- function(df) {
  # your code
  if(nrow(df) > 0 && ncol(df) == 0) return(0)
  if(ncol(df) == 0 | nrow(df) == 0) return(numeric(0))
  x<-apply(df, 2, is.na)
apply(x,1,sum)
}
```

```{r}
df1 <- read.csv(text = "  a,  b,  c
 10, NA, 10
 NA,  9, NA
  1,  2,  3
  4,  5,  6
", colClasses = "numeric")
 
# Expected:
#> [1] 1 2 0 0
countMissings(df1)
```
```{r}
# Expected:
#> numeric(0)
countMissings(data.frame())
```
```{r}
# Expected:
#> 1 2 3 4 5 6 
#> 0 0 0 0 0 0 
# or
#> [1] 0 0 0 0 0 0
countMissings(head(iris))
```


#### Tests
```{r, echo = FALSE, error = TRUE}
df1 <- read.csv(text = "  a,  b,  c
 10, NA, 10
 NA,  9, NA
  1,  2,  3
  4,  5,  6
", colClasses = "numeric")
testthat::expect_equal(countMissings(df1), c(1, 2, 0, 0))

testthat::expect_equal(unname(countMissings(head(iris))), rep(0, 6))

testthat::expect_equal(countMissings(data.frame(x = NA, y = c(1:3, NA))), c(1, 1, 1, 2))

testthat::expect_equal(countMissings(data.frame(a = 1)[0, , drop = FALSE]), numeric(0))

testthat::expect_equal(countMissings(data.frame(a = 1)[, 0, drop = FALSE]), 0)

testthat::expect_equal(countMissings(data.frame(x = NA, y = c(letters[1:3], NA))), c(1, 1, 1, 2))

cat("If there are no errors here then tests are ok.\n")
```



# Exercise 3: Remove Missing Cols

Write a faster version of this function which removes all columns from a data.frame that have missing values.
```{r}
dropMissingCols <- function(df) {
  drop <- character(0)
  for (col in colnames(df)) {
    for (row in seq_len(nrow(df))) {
      if (is.na(df[row, col])) drop <- union(drop, col)
    }
  }
  df[, setdiff(colnames(df), drop), drop = FALSE]
}
```

Expected Behaviour:
```{r}
testdf <- data.frame(x = c(1, 2, 3), y = c("A", "B", NA), z = c(1, NA, 2))
print(dropMissingCols(testdf))  # data.frame with all coluns that do not contain missings: 'x', in this case
```

#### Tests

```{r, echo = FALSE, error = TRUE}
dropMissingColsFast <- function(df) {
  df[colSums(is.na(df)) == 0]
}

# The following is necessary because the test server does not like loops.
i <- 1
  testdf <- iris
  testdf[cbind(sample(nrow(iris), i, replace = TRUE), sample(ncol(iris), i, replace = TRUE))] <- NA
  testthat::expect_equal(dropMissingCols(testdf), dropMissingColsFast(testdf))
i <- 2
  testdf <- iris
  testdf[cbind(sample(nrow(iris), i, replace = TRUE), sample(ncol(iris), i, replace = TRUE))] <- NA
  testthat::expect_equal(dropMissingCols(testdf), dropMissingColsFast(testdf))
i <- 3
  testdf <- iris
  testdf[cbind(sample(nrow(iris), i, replace = TRUE), sample(ncol(iris), i, replace = TRUE))] <- NA
  testthat::expect_equal(dropMissingCols(testdf), dropMissingColsFast(testdf))
i <- 4
  testdf <- iris
  testdf[cbind(sample(nrow(iris), i, replace = TRUE), sample(ncol(iris), i, replace = TRUE))] <- NA
  testthat::expect_equal(dropMissingCols(testdf), dropMissingColsFast(testdf))
i <- 5
  testdf <- iris
  testdf[cbind(sample(nrow(iris), i, replace = TRUE), sample(ncol(iris), i, replace = TRUE))] <- NA
  testthat::expect_equal(dropMissingCols(testdf), dropMissingColsFast(testdf))
i <- 6
  testdf <- iris
  testdf[cbind(sample(nrow(iris), i, replace = TRUE), sample(ncol(iris), i, replace = TRUE))] <- NA
  testthat::expect_equal(dropMissingCols(testdf), dropMissingColsFast(testdf))

```

# Exercise 4: Higher Order Fibonacci Sequence

Remember, the Fibonacci Sequence is defined as `F(n) = F(n-1) + F(n-2)`.

Here we define `x`th order fibonacci as `F(n, x) = F(n-1, x) + F(n-2, x) + ... + F(n - x, x)`.

The fibonacci numbers are recovered for `x == 2`.

Furthermore, `F(n, x)` is 1 for `n == 1`, and 0 for `n < 1`.

Write a faster version of this function:

```{r}
hoFib <- function(n, x) {
  if (n < 1) return(0)
  if (n == 1) return(1)
  result <- 0
  for (subtract in seq_len(x)) {
    result <- result + hoFib(n - subtract, x)
  }
  return(result)
}
```

Expected Behaviour:
```{r}
vapply(seq_len(10), function(n) hoFib(n, 3), numeric(1))  # c(1, 1, 2, 4, 7, 13, 24, 44, 81, 149)
```

#### Tests

```{r, echo = FALSE, error = TRUE}
hoFibFaster <- function(n, x) {
  if (n < 1) return(0)
  numbers <- numeric(n)
  numbers[[1]] <- 1
  for (i in seq_len(n - 1)) {
    sumstart <- i - x + 1
    if (sumstart < 1) sumstart <- 1
    numbers[[i + 1]] <- sum(numbers[seq.int(sumstart, i)])
  }
  return(numbers[[n]])
}
# The following is necessary because the test server does not like loops.

evalFib <- function(fibfun, top, x) sapply(seq_len(top), function(n) fibfun(n, x))

testthat::expect_equal(evalFib(hoFib, 15, 4), evalFib(hoFibFaster, 15, 4))
testthat::expect_equal(evalFib(hoFib, 15, 5), evalFib(hoFibFaster, 15, 5))
testthat::expect_equal(evalFib(hoFib, 15, 6), evalFib(hoFibFaster, 15, 6))
testthat::expect_equal(evalFib(hoFib, 15, 7), evalFib(hoFibFaster, 15, 7))
```

# Exercise 5: Balanced Parentheses

The following function checks if a given string of parentheses and square brackets is balanced.

The parentheses / brackets are balanced if for every opening paren, there is a corresponding closing paren; likewise for brackets.

```{r}
isBalanced <- function(string) {
  items <- strsplit(string, "")[[1]]
  stack <- character(0)
  is.balanced <- TRUE
  for (char in items) {
    if (char == "(") {
      stack <- c(stack, char)
    }
    if (char == "[") {
      stack <- c(stack, "[")
    }
    if (char == ")") {
      if (length(stack) == 0) {
        is.balanced <- FALSE
      } else {
        if (tail(stack, 1) != "(") is.balanced <- FALSE
        stack <- stack[seq_len(length(stack) - 1)]
      }
    }
    if (char == "]") {
      if (length(stack) == 0) {
        is.balanced <- FALSE
      } else {
        if (tail(stack, 1) != "[") is.balanced <- FALSE
        stack <- stack[seq_len(length(stack) - 1)]
      }
    }
  }
  is.balanced
}
```

Expected behaviour:
```{r}
isBalanced("")  # TRUE
isBalanced("()")  # TRUE
isBalanced("([])")  # TRUE
isBalanced("([)]")  # FALSE
isBalanced("()[)")  # FALSE
isBalanced("]()[")  # FALSE
```

#### Tests

```{r, echo = FALSE, error = TRUE}
testthat::expect_true(isBalanced(""))
testthat::expect_true(isBalanced("()"))
testthat::expect_true(isBalanced("([])"))
testthat::expect_true(!isBalanced("([)]"))
testthat::expect_true(!isBalanced("()[)"))
testthat::expect_true(!isBalanced("]()["))

testthat::expect_true(!isBalanced("]()["))

testthat::expect_false(isBalanced(")]]][[[[[]"))
testthat::expect_false(isBalanced("))[((]]])]](])[[)(()[[]][[[()())(])](][)][[[][))([[]]))](]]]([[[[(][[)()[)[())]()([[[]]]()])(([][]()"))
testthat::expect_true(isBalanced("[((()))][]([[]])(()[[]][[[()()]]])"))
testthat::expect_true(isBalanced("[((()))][]([[]])(()[[]][[[()()]]])[((()))][]([[]])(()[[]][[[()()]]])[((()))][]([[]])(()[[]][[[()()]]])"))
testthat::expect_true(isBalanced("([((()))][]([[]])(()[[]][[[()()]]])[((()))][]([[]])(()[[]][[[()()]]])[((()))][]([[]])(()[[]][[[()()]]]))"))
testthat::expect_true(!isBalanced("([((()))][]([[]])(()[[]][[[()()]]])[((()))][]([[]])(()[[]][[[()()]]])[((()))][]([[]])(()[[]][[[()()]]])]"))

```
