---
title: 'Lecture 05: Regular Expressions'
output: html_document
---


```{r setup, echo = FALSE}
## setup
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", out.width = "50%", echo = TRUE)
invisible(suppressPackageStartupMessages(testthat::expect_equal))
```

# Exercise 0: Test things

Define a variable `zero` with value `0`.
This is to try out the setup.
```{r}
zero <- 0  # change this?
```

# Exercise 1: Words

Define a variable `pattern.word` that matches words.
A word is a sequence of one or more `[[:alpha:]]` characters, surrounded by word boundaries `"\\b"`.

```{r}
pattern.word <- "\\b[A-z]+\\b"  # change this?
```

Expected behaviour:

```{r, error = TRUE}
some.words <- "word a7 I 1one, aNoThEr X!"

matched <- regmatches(some.words, gregexpr(pattern.word, some.words))[[1]]

matched # expected to be c("word", "I", "aNoThEr", "X")
```


#### Tests

```{r, echo = FALSE, error = TRUE}

some.words <- "word a7 I 1one,aNoThEr X!"

matched <- regmatches(some.words, gregexpr(pattern.word, some.words))[[1]]

testthat::expect_equal(matched, c("word", "I", "aNoThEr", "X"))

set.seed(1)
words.random <- replicate(10, paste(sample(c(letters, LETTERS), rpois(1, 5) + 1, replace = TRUE), collapse = ""))

nonwords.random <- replicate(10, paste(sample(c(letters, LETTERS, 0:9), rpois(1, 5) + 1, replace = TRUE), collapse = ""))

nonwords.random.true <- Filter(function(x) !all(strsplit(x, "")[[1]] %in% c(letters, LETTERS)), nonwords.random)


string <- paste(t(matrix(c(words.random, nonwords.random, sample(strsplit("!@#$%^&*(                ", "")[[1]], 20, replace = TRUE)), ncol = 2)), collapse = "")

matched <- regmatches(string, gregexpr(pattern.word, string))[[1]]

testthat::expect_equal(unique(sort(!!matched)), unique(sort(c(words.random, setdiff(nonwords.random, nonwords.random.true)))))
```

# Exercise 2: Even Words

Define a variable `pattern.word.even` that matches words of even length, i.e. words made up of an even number of characters.

```{r}
pattern.word.even <- "\\b([A-Za-z]{2})+\\b"  # change this?
```

Expected behaviour:

```{r, error = TRUE}
some.more.words <- "word a7 I 1one, aNoThEr Xx!"

matched <- regmatches(some.more.words, gregexpr(pattern.word.even, some.more.words))[[1]]

matched # expected to be c("word", "Xx")
```


#### Tests

```{r, echo = FALSE, error = TRUE}

some.more.words <- "word a7 I 1one, aNoThEr Xx!"

matched <- regmatches(some.more.words, gregexpr(pattern.word.even, some.more.words))[[1]]

testthat::expect_equal(matched, c("word", "Xx"))

set.seed(1)
words.random <- replicate(10, paste(sample(c(letters, LETTERS), rpois(1, 3) * 2 + 2, replace = TRUE), collapse = ""))

nonwords.random <- replicate(10, paste(sample(c(letters, LETTERS, 0:9), rpois(1, 2) * 2 + 2, replace = TRUE), collapse = ""))

nonwords.random.2 <- replicate(10, paste(sample(c(letters, LETTERS), rpois(1, 3) * 2 + 1, replace = TRUE), collapse = ""))

nonwords.random.true <- Filter(function(x) !all(strsplit(x, "")[[1]] %in% c(letters, LETTERS)), nonwords.random)


string <- paste(t(matrix(c(words.random, nonwords.random, nonwords.random.2,
  sample(strsplit("!@#$%^&*(                ", "")[[1]], 30, replace = TRUE)), ncol = 2)), collapse = "")

matched <- regmatches(string, gregexpr(pattern.word.even, string))[[1]]

testthat::expect_equal(unique(sort(!!matched)), unique(sort(c(words.random, setdiff(nonwords.random, nonwords.random.true)))))

```

# Exercise 3: Surrounded Words

Define a variable `pattern.word.surrounded` that matches words of length at least two letters where the first letter equals the last letter.
You probably want to use back-references (`"\\1"`) here.

```{r}
pattern.word.surrounded <- "\\b([A-z])([A-z]*)\\1\\b"  # change this?
```

Expected behaviour:

```{r, error = TRUE}
some.more.words.2 <- "wordw a7 I II 1one, aNoThEr Xx gling!"

matched <- regmatches(some.more.words.2, gregexpr(pattern.word.surrounded, some.more.words.2))[[1]]

matched # expected to be c("wordw", "II", "gling")
```

#### Tests

```{r, echo = FALSE, error = TRUE}

some.more.words.2 <- "wordw a7 I II 1one, aNoThEr Xx gling!"

matched <- regmatches(some.more.words.2, gregexpr(pattern.word.surrounded, some.more.words.2))[[1]]

testthat::expect_equal(matched, c("wordw", "II", "gling"))

set.seed(1)
words.random <- replicate(10, paste(sample(c(letters, LETTERS), rpois(1, 4) * 1, replace = TRUE), collapse = ""))
words.random <- paste0(words.random, substr(words.random, 1, 1))

nonwords.random <- replicate(10, paste(sample(c(letters, LETTERS, 0:9), rpois(1, 4) + 2, replace = TRUE), collapse = ""))

nonwords.random.2 <- replicate(10, paste(sample(c(letters, LETTERS), rpois(1, 4) + 2, replace = TRUE), collapse = ""))

nonwords.random.true <- Filter(function(x) {
  lets <- strsplit(x, "")[[1]]
  !all(lets %in% c(letters, LETTERS)) || lets[[1]] != tail(lets, 1)
}, c(nonwords.random, nonwords.random.2))


string <- paste(t(matrix(c(words.random, nonwords.random, nonwords.random.2,
  sample(strsplit("!@#$%^&*(                ", "")[[1]], 30, replace = TRUE)), ncol = 2)), collapse = "")

matched <- regmatches(string, gregexpr(pattern.word.surrounded, string))[[1]]

testthat::expect_equal(unique(sort(!!matched)), unique(sort(c(words.random, setdiff(c(nonwords.random.2, nonwords.random), nonwords.random.true)))))

```

# Exercise 4: Logical Scalar

Define a variable `pattern.logical` that matches a string if either `"TRUE"` or `"FALSE"`, but not otherwise.
Use the `^` and `$` to make sure your regex refers to the entire string.

```{r}
pattern.logical <- "^(TRUE|FALSE)$"  # change this?
```

Expected behaviour:

```{r, error = TRUE}
some.words.logical <- c("TRUE", "FALSE", "notTRUE", "FALSEx", "F.A.L.S.E", "", "TRUE")

matched <- grep(pattern.logical, some.words.logical, value = TRUE)

matched # expected to be c("TRUE", "FALSE", "TRUE")
```

#### Tests

```{r, echo = FALSE, error = TRUE}

some.words.logical <- c("TRUE", "FALSE", "notTRUE", "FALSEx", "F.A.L.S.E", "", "TRUE")

matched <- grep(pattern.logical, some.words.logical, value = TRUE)

testthat::expect_equal(matched, c("TRUE", "FALSE", "TRUE"))
```

# Exercise 5: Logical Vector

Define a variable `pattern.logical.vect` that matches a string if it is a representation of a logical vector, of the form
`"c(TRUE,FALSE,FALSE,TRUE,...)"`.

```{r}
pattern.logical.vect <- "^c\\((TRUE|FALSE)(,(TRUE|FALSE))*\\)$"  # change this?
```


Expected behaviour:

```{r, error = TRUE}
some.words.logical.vect <- c(
  "TRUE",  # do not accept single TRUE or FALSE, all matches should be in c().
  "FALSE",
  "c()",  # do not accept empty c()
  "",
  "c(TRUE)", # this should be accepted
  "c(FALSE)",
  "xc(FALSE)",  # not accepted: bad 'x' in the beginning
  "c(FASLE)",  # typo
  "c(TRUE,TRUE,FALSE)",
  "c(TRUE,TRUE,TRUE,TRUE)",
  "c(TRUE, TRUE)"  # do not accept spaces
)

matched <- grep(pattern.logical.vect, some.words.logical.vect, value = TRUE)

matched # expected to be c("c(TRUE)", "c(FALSE)", "c(TRUE,TRUE,FALSE)", "c(TRUE,TRUE,TRUE,TRUE)")
```

#### Tests

```{r, echo = FALSE, error = TRUE}
some.words.logical.vect <- c(
  "TRUE",  # do not accept single TRUE or FALSE, all matches should be in c().
  "FALSE",
  "c()",  # do not accept empty c()
  "",
  "c(TRUE)", # this should be accepted
  "c(FALSE)",
  "c(FALSE)x",
  "c(TUER)",
  "Ac(FALSE)",
  "c(TRUE,TRUE,FALSE)",
  "c(TRUE,TRUE,TRUE,TRUE)",
  "c(TRUE, TRUE)"  # do not accept spaces
)

matched <- grep(pattern.logical.vect, some.words.logical.vect, value = TRUE)

testthat::expect_equal(matched, c("c(TRUE)", "c(FALSE)", "c(TRUE,TRUE,FALSE)", "c(TRUE,TRUE,TRUE,TRUE)"))

set.seed(1)
nonwords.random <- replicate(10, paste(c("c(", sample(c(LETTERS, ",", ",", ","), rpois(1, 3) * 2 + 2, replace = TRUE), ")"), collapse = "", sep = ""))

words.random <- replicate(10, paste(c("c(", head(c(t(cbind(sample(c("TRUE", "FALSE"), rpois(1, 2) + 1, replace = TRUE), ","))), -1), ")"), collapse = "", sep = ""))

matched <- grep(pattern.logical.vect, c(words.random, nonwords.random))

testthat::expect_equal(matched, 1:10)
```

# Exercise 6: Prime

Define a variable `pattern.prime` that matches a string if it represents a prime number with up to three digits.

Hint 1: the prime numbers between 1 and 1000 are:
```{r}
Filter(function(x) sum(!x%%(2:x))<2, 2:1000)
```

Hint 2: It is probably easiest to write code that constructs this regular expression.

```{r}
pattern.prime <- paste0("\\b(", paste(Filter(function(x) sum(!x%%(2:x))<2, 2:1000), collapse = "|"), ")\\b")
```

Expected behaviour:
```{r, error = TRUE}
grep(pattern.prime, c("1", "2", "3", "4", "x3", "4x", "997", "995"), value = TRUE)  # only "2", "3", and "997".

# The following should give all prime numbers, i.e. as in `Filter(function(x) sum(!x%%(2:x))<2, 1:1000)`.
# Think about it: why does this work?
grep(pattern.prime, 1:10000)
```


#### Tests

```{r, echo = FALSE, error = TRUE}

primes <- Filter(function(x) sum(!x%%(2:x))<2, 2:1000)

testthat::expect_equal(grep(pattern.prime, c("1", "2", "3", "4", "x3", "4x", "997"), value = TRUE), c("2", "3", "997"))

testthat::expect_equal(grep(pattern.prime, 1:10000), primes)
```

# Exercise 7: Author Shortening

In some citation styles, author names are listed in the form `<last name>, <first initial>`, such as `Einstein, A.`.

Define variables `author.pattern` and `author.replacement` that can be used inside `sub()` (as in the examples to follow), so that the author format is changed from
"<first name> <last name>` to `<last name>, <first initial>`.
You likely need to use references in your replacement variable.
You can assume that first names always consist of only letters, while last names may contain spaces.
There are no malformed strings (i.e. all given values conform to this pattern), and no middle names.


```{r}
author.pattern <- "^([A-z]+)\\s+([A-z\\s]+)$"  # change this?
author.replacement <- "\\2, \\1{1}."  # change this?
```

```{r, error = TRUE}
words.names <- c("Klaus Dieter", "Johannes von Aalen", "Ou Wu")
sub(author.pattern, author.replacement, words.names)  # c("Dieter, K.", "von Aalen, J.", "Wu, O.")
```

#### Tests

```{r, echo = FALSE, error = TRUE}
words.names <- c("Klaus Dieter", "Johannes von Aalen", "Ou Wu")
testthat::expect_equal(sub(author.pattern, author.replacement, words.names), c("Dieter, K.", "von Aalen, J.", "Wu, O."))

set.seed(1)
first.random <- paste0(sample(LETTERS, 10, replace = TRUE), replicate(10, paste(sample(c(letters), rpois(1, 5) + 1, replace = TRUE), collapse = "")))
last.random <- replicate(10, gsub("  ", " ", paste(c(sample(LETTERS, 1), sample(c(letters, rep(" ", 5)), rpois(1, 5) + 1, replace = TRUE), sample(letters, 1)), collapse = "")))

names.in <- sprintf("%s %s", first.random, last.random)
names.expected <- sprintf("%s, %s.", last.random, substr(first.random, 1, 1))

testthat::expect_equal(sub(author.pattern, author.replacement, names.in), names.expected)



```
