---
title: 'Lecture: R6'
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, echo = FALSE}
## setup
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", out.width = "50%", echo = TRUE)
```

# R6

Solve the following tasks by creating or completing R6 classes.

## Exercise 01: Cat Herd (Cloning Behavior)

Create two classes, `Cat` and `CatHerd`, that illustrate deep cloning in R6.
The `CatHerd` class should contain a list of `Cat` objects and implement the `deep_clone()` method such that modifications to the original herd do not affect the cloned herd.

```{r}
Cat <- R6::R6Class("Cat",
  public = list(
    name = NULL,
    age = NULL,
    initialize = function(name, age) {
      self$name <- name
      self$age <- age
    }
  )
)

CatHerd <- R6::R6Class("CatHerd",
  public = list(
    cats = NULL,
    initialize = function(cats = list()) {
      self$cats <- cats
    }
  ),
  private = list(
    # implement deep_clone
           deep_clone = function(name, value) {
             #name must be equal to just like in the public fields
                           if (name == "cats") {
                             # Clone each cat object in the list
                             # pay attention that we used lowered case cat
                             lapply(value, function(cat) cat$clone(deep = TRUE))
                           } else {
                             # For other fields, return the value as is
                             value
                           }
                         }
  )
)
```

### Example Behviour:

```{r}
cat1 <- Cat$new("Fluffy", 2)
cat2 <- Cat$new("Whiskers", 3)
herd <- CatHerd$new(list(cat1, cat2))

# Deep clone
herd.clone <- herd$clone(deep = TRUE)

# Modify an element in the original herd
herd$cats[[1]]$age <- 4

# Check that herd changed, but herd.clone is not affected
herd$cats[[1]]$age
## Should be:
#> [1] 4
herd.clone$cats[[1]]$age
## Should be:
#> [1] 2
```

Running the above code should show that **only** the age in the original herd is changed to 4, while the cloned herd’s first cat remains age 2.

### Tests

```{r, echo = FALSE, error = TRUE}
cat1 <- Cat$new("Fluffy", 2)
cat2 <- Cat$new("Whiskers", 3)
herd <- CatHerd$new(list(cat1, cat2))
herd.clone <- herd$clone(deep = TRUE)
checkmate::expect_r6(herd, "CatHerd")
checkmate::expect_r6(herd.clone, "CatHerd")

testthat::expect_equal(herd, herd.clone)
herd$cats[[1]]$age <- 4
testthat::expect_equal(herd$cats[[1]]$age, 4)
testthat::expect_equal(herd.clone$cats[[1]]$age, 2)
testthat::expect_equal(herd$cats[[2]]$age, 3)
testthat::expect_equal(herd.clone$cats[[2]]$age, 3)
herd$cats[[2]]$age <- 4
testthat::expect_equal(herd$cats[[2]]$age, 4)
testthat::expect_equal(herd.clone$cats[[2]]$age, 3)

# Check that the cat objects inside each herd are separate
testthat::expect_false(identical(herd$cats[[1]], herd.clone$cats[[1]]))

cat("If there are no errors here then tests are ok.\n")
```

## Exercise 02: Rectangle with Active Binding

Create a class `Rectangle` with two public fields `width` and `height`.
It should be possible to set these on construction.
Implement an active binding `area` that computes the rectangle’s area dynamically (i.e., updates automatically when `width` or `height` change).
It should be able to assign a value to `area`, in which case the `width` should be adjusted accordingly.

```{r}
Rectangle <- R6::R6Class("Rectangle",
  public = list(
    width = NULL,
    height = NULL,
    initialize = function(width, height) {
      # your code
        self$width <- width
      self$height <- height
    }
  ),
  active = list(
    # your code
    # This is in the case that you want to dynamically uptade some variable
    area = function(value) {
      if (missing(value)) {
        # Getter: calculate and return the area dynamically
        return(self$width * self$height)
      } else {
        # Setter: adjust width to match the new area, keeping height constant
        #essentialy 'value' is the new area 
        #And you want to find the width that can be multiplied with height to get
        #that area
        self$width <- value / self$height
      }
    }
  )
)
```

### Example Behaviour

```{r , error= TRUE}
rect <- Rectangle$new(width = 5, height = 10)
rect$area
## Should be:
#> [1] 50

rect$width <- 6
rect$area
## Should be:
#> [1] 60

rect$height <- 3
rect$area
## Should be:
#> [1] 18

rect$area <- 27
rect$area
## Should be
#> [1] 27
rect$height
## Should be
#> [1] 3
rect$width
## Should be
#> [1] 9
```

Running this code should confirm that the `area` updates whenever either `width` or `height` is changed.
Attempting to do `rect$area <- 100` should result in an error.

### Tests

```{r, echo = FALSE, error = TRUE}
# Test setup
rect.test <- Rectangle$new(width = 5, height = 10)

# Check class
checkmate::expect_r6(rect.test, "Rectangle")

# Check initial area
testthat::expect_equal(rect.test$area, 50)

# Update width
rect.test$width <- 6
testthat::expect_equal(rect.test$area, 60)

# Update height
rect.test$height <- 3
testthat::expect_equal(rect.test$area, 18)

rect.test$area <- 27

testthat::expect_equal(rect.test$area, 27)
testthat::expect_equal(rect.test$width, 9)
testthat::expect_equal(rect.test$height, 3)

cat("If there are no errors here then tests are ok.\n")
```

**Task**: Implement the `area` active binding in `Rectangle` so that it always returns `width * heigth` and throws an error if the user tries to assign to `area`.
```{r}
Rectangle <- R6::R6Class("Rectangle",
  public = list(
    width = NULL,
    height = NULL,
    initialize = function(width, height) {
      # your code
        self$width <- width
      self$height <- height
    }
  ),
  active = list(
    # your code
    # This is in the case that you want to dynamically uptade some variable
    area = function(value) {
      if (missing(value)) {
        # Getter: calculate and return the area dynamically
        return(self$width * self$height)
      } else {
       stop("cannot assign new area")
      }
    }
  )
)
```
## Exercise 03: Bank Account (`private`)

Write a class `BankAccount` that has two functions, `$deposit()` and `$withdraw()`, both with single numeric argument `amount`.
For both, the argument should be checked to be a positive scalar numeric.

The class should have a *private* field `.balance` (note the dot), and the balance should be adjusted by `$deposit()` and `$withdraw()`.
Negative balances should be prohibited: Reducing the balance to a negative amount should throw an error before the class is changed.
Both `deposit()` and `$withdraw()` should return `invisible(self)`.

The constructor should have an argument `initial.amount`, which should be the starting value of `.balance`.

There should be an active binding `$balance`, which should return the current balance.

```{r}
library("checkmate")
BankAccount <- R6::R6Class("BankAccount",
                           public = list(
                             # your code...
                             initial.amount = NULL,
                             amount = NULL,
                             initialize = function(initial.amount) {
                               private$.balance <- initial.amount
                               invisible(self)
                             },
                             withdraw = function(amount) {
                               # your code
                               if(private$.balance < amount) {
                                 stop("Insufficient funds.")
                               }
                               assertNumber(amount,lower = 1)
                               private$.balance <-  private$.balance - amount
                               invisible(self)
                             },
                             deposit = function(amount) {
                               # your code
                               assertNumber(amount,lower = 1)
                               private$.balance <-  private$.balance + amount
                               invisible(self)
                             }
                           ),
                           active = list(
                             # your code
                             balance = function(value) {
                               if (missing(value)) {
                                 private$.balance
                               } else {
                                 stop("You cant set your own balance")
                               }
                               
                             }
                           ),
                           private = list(
                             # your code
                             .balance = NULL
                           )
)

```

Expected behavior:

```{r, error = TRUE}
account.georg <- BankAccount$new(10)

account.georg$balance
# Expected:
#> [1] 10

account.kunigunde <- BankAccount$new(20)
account.kunigunde$deposit(10)$withdraw(5)

account.kunigunde$balance
# Expected:
#> [1] 25

try(account.kunigunde$balance <- 1000)  # should give an error

try(account.kunigunde$withdraw(1000))  # should give an error
```

### Tests

```{r, echo = FALSE, error = TRUE}
account.georg <- BankAccount$new(10)

testthat::expect_equal(account.georg$balance, 10)

account.kunigunde <- BankAccount$new(20)
testthat::expect_identical(account.kunigunde$deposit(10)$withdraw(5), account.kunigunde)

testthat::expect_equal(account.kunigunde$balance, 25)
testthat::expect_error(account.kunigunde$balance <- 1000)  # should give an error

testthat::expect_error(account.kunigunde$withdraw(1000))  # should give an error

testthat::expect_equal(names(account.kunigunde$.__enclos_env__$private), ".balance")

cat("If there are no errors here then tests are ok.\n")
```
