---
title: 'Unit 04: Structured Programming III'
output: html_document
---

```{r setup, echo = FALSE}
## setup
library("checkmate")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", out.width = "50%", echo = TRUE)
```

# Checkmate, Lapply

## Setup

Make sure you run this to load the checkmate package.

```{r}
library("checkmate")
```

## Exercise 0: Test things

Define a variable `zero` with value `0`.
This is to get used to the setup.
```{r}
zero <- 1  # change this?
```

## Exercise 1: checkmate

Use the correct checkmate functions:
Fill in the correct `assert` for the given argument from the description.

### Exercise 1a

`n`: a scalar numeric, no missing values.

```{r}
f1a <- function(n) {
  # your code 
}
```

#### Tests

```{r, echo = FALSE, error = TRUE}
invisible(f1a(1))
testthat::expect_error(f1a("a"))
testthat::expect_error(f1a(list(1)))
testthat::expect_error(f1a(c(1, 1)))
testthat::expect_error(f1a(NA_real_))
testthat::expect_error(f1a(NULL))
testthat::expect_error(f1a())
cat("If there are no errors here then tests are ok.\n")
```

### Exercise 1b

`s`: a character vector with at least 2 elements; no missing values

```{r}
f1b <- function(s) {
  # your code 
}
```


#### Tests
```{r, echo = FALSE, error = TRUE}
# Valid test case
invisible(f1b(c("apple", "banana")))

# Invalid cases
testthat::expect_error(f1b("apple"))               # Only one element
testthat::expect_error(f1b(c("apple", NA)))        # Contains missing value
testthat::expect_error(f1b(c("apple", "banana", NA))) # Contains missing value
testthat::expect_error(f1b(1:3))                   # Not a character vector
testthat::expect_error(f1b(list("apple", "banana"))) # List instead of character vector
testthat::expect_error(f1b(NULL))                  # NULL input
testthat::expect_error(f1b())                      # No argument provided
cat("If there are no errors here then tests are ok.\n")
```

### Exercise 1c

`i`: scalar integer-valued numeric, missing values are allowed.

```{r}
f1c <- function(i) {
  # your code 
}
```

#### Tests
```{r, echo = FALSE, error = TRUE}
# Valid test cases
invisible(f1c(1))                # Single integer
invisible(f1c(NA_integer_))      # NA integer
invisible(f1c(3L))               # Explicit integer type

# Invalid cases
testthat::expect_error(f1c("a"))           # Non-numeric input
testthat::expect_error(f1c(1.5))           # Non-integer numeric (decimal)
testthat::expect_error(f1c(c(1, 2)))       # Vector instead of scalar
testthat::expect_error(f1c(NULL))          # NULL input
testthat::expect_error(f1c(list(1)))       # List instead of scalar
testthat::expect_error(f1c())              # No argument provided
cat("If there are no errors here then tests are ok.\n")
```

### Exercise 1d

`v`: A scalar character, either `"yes"`, `"no"`, or `"maybe"`.

```{r}
f1d <- function(v) {
  # your code 
}
```

#### Tests
```{r, echo = FALSE, error = TRUE}
# Valid test cases
invisible(f1d("yes"))
invisible(f1d("no"))
invisible(f1d("maybe"))

# Invalid cases
testthat::expect_error(f1d("perhaps"))      # Invalid choice
testthat::expect_error(f1d("YES"))          # Case-sensitive mismatch
testthat::expect_error(f1d(c("yes", "no"))) # Vector instead of scalar
testthat::expect_error(f1d(1))              # Non-character input
testthat::expect_error(f1d(NA_character_))  # NA value
testthat::expect_error(f1d(NULL))           # NULL input
testthat::expect_error(f1d())               # No argument provided
cat("If there are no errors here then tests are ok.\n")
```
### Exercise 1e

`m`: Either a `matrix`, or a `data.frame`, having at least two rows, at least two columns, and containing only numeric, non-missing values.

```{r}
f1e <- function(m) {
  # your code 
}
```

#### Tests
```{r, echo = FALSE, error = TRUE}
# Valid test cases
invisible(f1e(matrix(1:4, nrow = 2, ncol = 2)))               # 2x2 numeric matrix
invisible(f1e(data.frame(a = c(1, 2), b = c(3, 4))))          # 2x2 numeric data frame
invisible(f1e(matrix(1:9, nrow = 3, ncol = 3)))               # 3x3 numeric matrix
invisible(f1e(data.frame(a = c(1, 2, 3), b = c(4, 5, 6))))    # 3x2 numeric data frame

# Invalid cases
testthat::expect_error(f1e(matrix(1:4, nrow = 1, ncol = 4)))        # Less than 2 rows
testthat::expect_error(f1e(matrix(1:4, nrow = 4, ncol = 1)))        # Less than 2 columns
testthat::expect_error(f1e(matrix(c(1, 2, NA, 4), nrow = 2)))      # Contains missing value
testthat::expect_error(f1e(data.frame(a = c(1, NA), b = c(3, 4)))) # Missing value in data frame
testthat::expect_error(f1e(data.frame(a = c("a", "b"), b = c("c", "d")))) # Non-numeric data frame
testthat::expect_error(f1e(matrix(c("a", "b", "c", "d"), nrow = 2))) # Non-numeric matrix
testthat::expect_error(f1e(data.frame(a = 1:2, b = c(3.5, "test")))) # Mixed types in data frame
testthat::expect_error(f1e(1:4))                                    # Not a matrix or data frame
testthat::expect_error(f1e(list(matrix(1:4, nrow = 2, ncol = 2))))   # Wrapped in list instead of matrix/data frame
testthat::expect_error(f1e(NULL))                                    # NULL input
testthat::expect_error(f1e())                                        # No argument provided

cat("If there are no errors here then tests are ok.\n")
```


## Exercise 2: Loops, lapply etc.

### Exercise 2a

Write a function that takes two arguments: `words` and `text`. Both are character vectors. The function should return a logical vector indicating, for each word in `words`, whether it occurs in `text`.

```{r}
wordInText <- function(words, text) {
  # your code
}
```

Example:

```{r}
text.example <- c("this", "is", "a", "small", "text")
words.example <- c("this", "not", "small")
```

Expected return:

```{r}
wordInText(words.example, text.example)
```
```r
#> [1] TRUE FALSE TRUE
```

#### Tests

```{r, echo = FALSE, error = TRUE}
text_example <- c("this", "is", "a", "small", "text")
words_example <- c("this", "not", "small")
result <- wordInText(words_example, text_example)
expected <- c(TRUE, FALSE, TRUE)
testthat::expect_equal(result, expected)

words_example <- c("a", "small", "text", "example")
result <- wordInText(words_example, text_example)
expected <- c(TRUE, TRUE, TRUE, FALSE)
testthat::expect_equal(result, expected)

text_example <- c("hello", "world")
words_example <- c("hello", "world", "test")
result <- wordInText(words_example, text_example)
expected <- c(TRUE, TRUE, FALSE)
testthat::expect_equal(result, expected)

text_example <- character(0)
words_example <- c("any", "word")
result <- wordInText(words_example, text_example)
expected <- c(FALSE, FALSE)
testthat::expect_equal(result, expected)

words_example <- character(0)
result <- wordInText(words_example, text_example)
expected <- logical(0)
testthat::expect_equal(result, expected)

cat("If there are no errors here then tests are ok.\n")
```

### Exercise 2b

Write a function that takes two arguments: `words` and `text`. Both are character vectors. The function should return a numeric vector indicating, for each word in `words`, how many times it occurs in `text`.

You may need to use `unname()` to pass the tests.

Example:

```{r}
countWordOccurrences <- function(words, text) {
  # your code
}
```

```{r}
text.example <- c("this", "is", "a", "small", "text", "this", "is", "a", "test")
words.example <- c("this", "a", "test", "not")
```

Expected return:

```{r}
countWordOccurrences(words.example, text.example)
```
```r
#> [1] 2 2 1 0
```


#### Tests

```{r, echo = FALSE, error = TRUE}
text_example <- c("this", "is", "a", "small", "text", "this", "is", "a", "test")
words_example <- c("this", "a", "test", "not")
result <- countWordOccurrences(words_example, text_example)
expected <- c(2, 2, 1, 0)
testthat::expect_equal(result, expected)

words_example <- c("small", "text", "is")
result <- countWordOccurrences(words_example, text_example)
expected <- c(1, 1, 2)
testthat::expect_equal(result, expected)

text_example <- c("hello", "world", "hello", "world", "hello")
words_example <- c("hello", "world", "test")
result <- countWordOccurrences(words_example, text_example)
expected <- c(3, 2, 0)
testthat::expect_equal(result, expected)

text_example <- character(0)
words_example <- c("any", "word")
result <- countWordOccurrences(words_example, text_example)
expected <- c(0, 0)
testthat::expect_equal(result, expected)

words_example <- character(0)
result <- countWordOccurrences(words_example, text_example)
expected <- numeric(0)
testthat::expect_equal(result, expected)

cat("If there are no errors here then tests are ok.\n")
```

---

### Exercise 2c

Write a function that takes a matrix as input and returns a numeric vector indicating, for each row, how many `NA` values it contains.


```{r}
countNAsPerRow <- function(mat) {
  # your code
}
```

Example:

```{r}
mat.example <- matrix(c(1, NA, 3, 4, 5, NA), nrow = 2)
```
```r
#>      [,1] [,2] [,3]
#> [1,]    1    3    5
#> [2,]   NA    4   NA
```

Expected return:

```{r}
countNAsPerRow(mat.example)
```
```r
#> [1] 0 2
```

#### Tests

```{r, echo = FALSE, error = TRUE}
mat_example <- matrix(c(1, NA, 3, 4, 5, NA), nrow = 2)
result <- countNAsPerRow(mat_example)
expected <- c(0, 2)
testthat::expect_equal(result, expected)

mat_example <- matrix(c(NA, NA, NA, 1, 2, 3), nrow = 3)
result <- countNAsPerRow(mat_example)
expected <- c(1, 1, 1)
testthat::expect_equal(result, expected)

mat_example <- matrix(1:9, nrow = 3)
result <- countNAsPerRow(mat_example)
expected <- c(0, 0, 0)
testthat::expect_equal(result, expected)

mat_example <- matrix(rep(NA, 9), nrow = 3)
result <- countNAsPerRow(mat_example)
expected <- c(3, 3, 3)
testthat::expect_equal(result, expected)

mat_example <- matrix(numeric(0), nrow = 0, ncol = 0)
result <- countNAsPerRow(mat_example)
expected <- numeric(0)
testthat::expect_equal(result, expected)

cat("If there are no errors here then tests are ok.\n")
```
