---
title: 'Lecture 02: Idiomatic R'
output: html_document
---


```{r setup, echo = FALSE}
## setup
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", out.width = "50%", echo = TRUE)
```

# Idiomatic R

Solve the following tasks.
See if you know specific `R` functions or tricks that these exercises particularly easy.
Try to solve these without loading any additional `R`-packages.


## Exercise 00: Zero

(Demo Exercise)

Create a variable `zero` that is a `numeric` scalar containin the number zero, i.e. `0`.
```{r}
# your code
zero<- 0
```

#### Tests
```{r, echo = FALSE, error = TRUE}
checkmate::expect_numeric(zero, lower = 0, upper = 0, any.missing = FALSE, len = 1)
cat("If there are no errors here then tests are ok.\n")
```

## Exercise 01: Zeroes

Create a variable `z` that is a `numeric` vector consisting of 1024 zeroes, i.e. `c(0, 0, 0, ...., 0)`.
```{r}
# your code
z <- rep(0,1024)
z
```

#### Tests
```{r, echo = FALSE, error = TRUE}
checkmate::expect_numeric(z, lower = 0, upper = 0, any.missing = FALSE, len = 1024)
cat("If there are no errors here then tests are ok.\n")
```

## Exercise 02: Powers of Ten

Create a variable `tens` that is a `numeric` vector of length 16 consisting of powers of ten in increasing order, starting at 1:
`c(1, 10, 100, 1000, ..... 10 ^ 15)`

```{r}
# your code
tens <- 10^seq(0,15)
tens
```

#### Tests
```{r, echo = FALSE, error = TRUE}
checkmate::expect_numeric(tens, lower = 1, upper = 10^19, any.missing = FALSE, len = 16, unique = TRUE, sorted = TRUE)
testthat::expect_true(all(tens[-1] %% 10 == 0))
testthat::expect_equal(tens[[1]], 1)
cat("If there are no errors here then tests are ok.\n")
```
## Exercise 03: Simple Function

Write a function that increments every second element of a numeric vector.
Inputs:

* **`x`**: vector with even number of elements, e.g. `c(1, 1)` or `c(-1, 0, 10, 9)`

Return value:
The value of **`x`** where every second element is increased by one.

```{r}
secondInc <- function(x) {
  # your code
  #x[c(FALSE,TRUE)] <- x[c(FALSE,TRUE)] +1
  x[seq(2,length(x),2)] <- x[seq(2,length(x),2)] +1

x
}
```

Example calls:
```{r}
secondInc(c(1, 1))
# Expected:
##> [1] 1 2
```
```{r}
secondInc(c(-1, 0, 10, 9))
# Expected:
##> [1] -1  1 10 10
```
```{r}
secondInc(c(1, -1, 1, -1, 1, -1))
# Expected:
##> [1] 1 0 1 0 1 0
```
```{r}
x <- secondInc(c(0, -1))
# This should *not* generate any output!
```
```{r}
x
# Expected
##> [1] 0 0
```

#### Tests
```{r, echo = FALSE, error = TRUE}
testthat::expect_equal(secondInc(c(1, 1)), c(1, 2))
testthat::expect_equal(secondInc(c(-1, 0, 10, 9)), c(-1, 1, 10, 10))
testthat::expect_equal(secondInc(c(1, -1, 1, -1, 1, -1)), c(1, 0, 1, 0, 1, 0))
testthat::expect_output(secondInc(c(1, 1)), NA)

testvector <- rpois(10, 5)
testthat::expect_equal(secondInc(c(secondInc(!!testvector)[-1], 0)), c(testvector[-1], 0) + 1)
cat("If there are no errors here then tests are ok.\n")
```

## Exercise 04

Write a function returns a greeting for the user.
Inputs:

* **`name`**: length-1 `character` vector

Return value:
The return value should be `"Hello, <name>"`.

Your function should not generate any output, only return the string!

```{r}
greet <- function(name) {
  # your code
paste("Hello,",name)}
```

Example calls:
```{r}
greet("Helmut")        # "Hello, Helmut"
greet("Anna Maier")    # "Hello, Anna Maier"
x <- greet("Sarah")    #  no output!
x                      # "Hello, Sarah"
```


#### Tests
```{r, echo = FALSE, error = TRUE}
testthat::expect_equal(greet("Helmut"), "Hello, Helmut")
testthat::expect_equal(greet("Anna Maier"), "Hello, Anna Maier")
testthat::expect_equal(greet("Sarah"), "Hello, Sarah")
testthat::expect_output(greet("Helmut"), NA)

st <- paste(sample(letters, 4), collapse = "")
checkmate::expect_string(greet(st), pattern = "^Hello, ")
checkmate::expect_string(greet(st), pattern = st)

cat("If there are no errors here then tests are ok.\n")
```

## Exercise 05

Write a function that greets the user a given number of times.
The function should output the line `Hello` a given number of times, using `cat()`.
Inputs:

* **`n`**: Number of greetings to print.

Return value:
The function should return the number of times it printed the greeting, i.e. the value of `n`.

```{r}
multiGreet <- function(n) {
for (i in seq_len(n)) {
    cat("Hello\n")
} 
  n
}
```

Example calls:
```{r}
multiGreet(3)
# Expected
##> Hello
##> Hello
##> Hello
##> [1] 3
```
Example calls:
```{r}
multiGreet(0)
# Expected
##> [1] 0
```

#### Tests
```{r, echo = FALSE, error = TRUE}
testthat::expect_output(testthat::expect_equal(multiGreet(3), 3), "^Hello\nHello\nHello$")
testthat::expect_output(testthat::expect_equal(multiGreet(1), 1), "^Hello$")
testthat::expect_output(testthat::expect_equal(multiGreet(0), 0), NA)
testthat::expect_output(testthat::expect_equal(multiGreet(5), 5), "^(Hello\n){4}Hello$")


cat("If there are no errors here then tests are ok.\n")
```

## Exercise 06

Write a function that takes a `numeric` vector and, for each element, calculates the difference to the value of the element coming two entries later.
E.g. `c(1, 2, 3, 4, 5)` --> `c(3 - 1, 4 - 2, 5 - 3)` --> `c(2, 2, 2)`.
`c(3, 30, 2, 20, 1, 10)` --> `c(2 - 3, 20 - 30, 1 - 2, 10 - 20)` --> `c(-1, -10, -1, -10)`.

Inputs:
* **`x`**: `numeric` vector, at least length 3, for which to calculate differences.

Return value: A vector that has length `length(x) - 2`, where the i'th element has the value `x[[i + 2]] - x[[i]]`.


```{r}
diffOverNext <- function(x) {
  # your code
  
result <- numeric(0)

for (i in seq_len(length(x) -2)) {
  result[[length(result) + 1]] <-  x[[i+2]] -x[[i]]
}
result
}
```
Example calls:
```{r}
diffOverNext(c(1, 2, 3, 4, 5))  # c(2, 2, 2)
diffOverNext(c(3, 30, 2, 20, 1, 10))  # c(-1, -10, -1, -10)
diffOverNext(c(1, 2, 3))  # 2
diffOverNext(c(1, 2, 1, 2, 1, 2))  # c(0, 0, 0, 0)
```

#### Tests
```{r, echo = FALSE, error = TRUE}
testthat::expect_equal(diffOverNext(c(1, 2, 3, 4, 5)), c(2, 2, 2))
testthat::expect_equal(diffOverNext(c(3, 30, 2, 20, 1, 10)), c(-1, -10, -1, -10))
testthat::expect_equal(diffOverNext(c(1, 2, 3)), 2)
testthat::expect_equal(diffOverNext(c(1, 2, 1, 2, 1, 2)), c(0, 0, 0, 0))

res1 <- rpois(4, 2)
res2 <- rpois(4, 3)


integ <- c(rbind(c(0, cumsum(res1)), c(0, cumsum(res2))))
expect <- c(rbind(res1, res2))
testthat::expect_equal(diffOverNext(integ), expect)

cat("If there are no errors here then tests are ok.\n")
```

## Exercise 07

Write a function that calculates the day of the year, given day and month, assuming the current year is not a leap-year (i.e. "kein Schaltjahr").
`"January 1"` is day `1`.
`"January 31"` is day `31`.
`"February 1"` is day `32`.
`"March 1"` is day `60`.
Etc.

Do this *without* using any date-related functions in `R`, do not use e.g. `as.Date()` or `strptime()`.

Inputs:

* **`month`**: `character(1)`: month of the year, written in english, starting with a capital letter (`"January"`, `"February"`, ...)
* **`day`**: `numeric(1)`: day of the year, scalar integer value between 1 and 31.

```{r}
dayNumber <- function(month, day) {
  # your code
  
months_list <- c("January" = 1, "February" = 2, "March" = 3, "April" = 4,
                 "May" = 5, "June" = 6, "July" = 7, "August" = 8,
                 "September" = 9, "October" = 10, "November" = 11, "December" = 12)

# Days in each month (assuming a non-leap year)
days_in_month <- c("January" = 31, "February" = 28, "March" = 31, "April" = 30,
                   "May" = 31, "June" = 30, "July" = 31, "August" = 31,
                   "September" = 30, "October" = 31, "November" = 30, "December" = 31)


n <- 1
result <- 0
repeat{
  if(month == names(days_in_month)[[n]]){
    break
  }
  result <- result +days_in_month[[n]]
  
  n <- n + 1
  
}
final <-result + day
final
}
```

Example calls:
```{r}
dayNumber("January", 1)   # 1
dayNumber("January", 31)  # 31
dayNumber("February", 1)  # 32
dayNumber("March", 1)     # 60
dayNumber("December", 31) # 365
```

#### Tests
```{r, echo = FALSE, error = TRUE}
testthat::expect_equal(dayNumber("January", 1), 1)
testthat::expect_equal(dayNumber("January", 2), 2)
testthat::expect_equal(dayNumber("February", 1), 32)
testthat::expect_equal(dayNumber("March", 1), 60)
testthat::expect_equal(dayNumber("December", 31), 365)

dn <- round(runif(1, 1, 31))
testthat::expect_equal(dayNumber("October", !!dn), 
  as.numeric(difftime(as.Date(sprintf("22-10-%s", dn)), as.Date("21-12-31"), units = "days"))
)

dn <- round(runif(1, 1, 30))
testthat::expect_equal(dayNumber("June", !!dn), 
  as.numeric(difftime(as.Date(sprintf("22-06-%s", dn)), as.Date("21-12-31"), units = "days"))
)

cat("If there are no errors here then tests are ok.\n")
```

## Exercise 08`*`

This exercise is more difficult than the preceding exercises.
We do not expect you to be able to solve it yet, but we also do not want you to get bored.

Write a function that determines the "nesting depth" of a list of lists.

The input consists only of lists that may or may not contain other lists, which again may contain lists etc.

* The nesting depth of `list()` is 0.
* The nesting depth of `list(list())` is 1.
* The nesting depth of `list(list(), list())` is 1.
* The nesting depth of `list(list(), list(list(), list()))` is 2.

--> The nesting depth of a list with elements `a`, `b`, `c`, .... is therefore
   1 + the maximum nesting depth of any of `a`, `b`, `c`, ... .

```{r}
nestingDepth <- function(lst) {
  # your code
  length(lst)
}
```

```{r}
nestingDepth(list())                      # 0
nestingDepth(list(list()))                # 1
nestingDepth(list(list(), list(list())))  # 2
nestingDepth(list(list(), list()))        # 1
```

#### Tests
```{r, echo = FALSE, error = TRUE}

testthat::expect_equal(nestingDepth(list()), 0)
testthat::expect_equal(nestingDepth(list(list())), 1)
testthat::expect_equal(nestingDepth(list(list(), list())), 1)
testthat::expect_equal(nestingDepth(list(list(), list(), list())), 1)
testthat::expect_equal(nestingDepth(list(list(), list(list()), list())), 2)

l <- list()
for (i in seq_len(4)) {
  l <- list(list(), list(l), list(list(l)))
  testthat::expect_equal(nestingDepth(l), 3 * i)
}

cat("If there are no errors here then tests are ok.\n")
```

